---
title: "Homwork 2"
author: "chenxin zhang cz2634"
date: "09/28/2020"
output: github_document
---

```{r setup, message = FALSE, echo = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1
#### Read the Mr.Trashwheel dataset.

```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```

```{r message = FALSE, echo = FALSE}
trashwhell_df = 
  read_xlsx(
            path = path_to_data,
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")) %>% 
            janitor::clean_names() %>% 
            drop_na(dumpster) %>% 
            mutate(
               sports_balls = round(sports_balls),
              sports_balls = as.integer(sports_balls)
    )
```

```{r}
trashwhell_df
```


#### Read precipitation data for 2018 and 2017.
** 2018
```{r echo = FALSE}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1) %>% #skip the first row of the sheet
  janitor::clean_names() %>%
  drop_na(month) %>% #drop the row based on the missing values in var. 'month'
  mutate(year = 2018) %>%
  relocate(year)
precip_2018
```
** 2017
```{r precip_2017, echo = FALSE}
precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1) %>% 
  janitor::clean_names() %>%
  drop_na(month) %>% 
  mutate(year = 2017) %>%
  relocate(year)
precip_2017
```

```{r echo = TRUE, collapse = TRUE}
# create a data frame for month
month_df = 
  tibble(
    month = 1:12, 
    month_name = month.name
  )
```

```{r message = FALSE, echo = FALSE}
# row-combine precip_2018 and precip_2017
precip_df = 
  bind_rows(precip_2018, precip_2017) 
precip_m = left_join(precip_df, month_df, by = "month") 
precip_m 
```


 This dataset contains information from the Mr.Trashwheel trash collector in Baltimore, Maryland. As trash enters the  inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwhell_df)` rows in our final dataset. Additional data sheets include monthly precipitation for year 2017 and 2018. The total monthly precipitation for year 2017 is `r sum(pull(precip_2017, total))` (in). The total monthly precipitation for year 2018 is `r sum(pull(precip_2018, total))` (in). 

## Problem 2

#### Read and Clean the NYC Transit dataset

```{r message = FALSE, echo = FALSE}
transit_df = 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%  # transfer to lowercase
  select(line:entry, vending, ada) %>% #keep specific columns
  mutate(entry = recode(entry, "Yes" = "True", "No" = "False"),
         entry = as.logical(entry)) # change 'entry' from charactor variable to logical variable
transit_df
```

The dataset is consist of `r nrow(transit_df)` obervations and `r ncol(transit_df)` varaibles. The variables include `r names(transit_df)`. Some works I did so far include:  
* clean the varaible names to lowercase, 
* choose columns that we are interested in, 
* recode a character varaible(entry) to a logical varaible.

* The NYC transits dataset includes `r transit_df %>% distinct(station_name, line) %>% nrow ` distinct stations. 
* Among these stations, `r transit_df %>% distinct(station_name, line, ada) %>% filter(ada == TRUE) %>% nrow ` stations are ADA compliant. 
* The proportion of station entrances/exits without vending allow entrance is `r (filter(transit_df, vending == "NO", entry =="TRUE") %>% nrow / filter(transit_df, vending == "NO") %>% nrow)*100`%.

#### Reformat data so that route number and route name are distinct variables.

```{r message = FALSE, echo = FALSE}
transit_tidy = 
  transit_df %>% 
  mutate_at(vars(route1:route11), as.character) %>% # convert number to character
  pivot_longer(
    route1:route11,
    names_to = "route_name",
    values_to = "route_number") %>% 
  drop_na(route_number)
transit_tidy
```

* The transit dataset shows there are `r transit_tidy %>% filter(route_number == "A") %>% distinct(station_name, line) %>% count()` stations serve the A train.
* Of the stations that serve the A train, `r transit_tidy %>% filter(route_number == "A") %>% distinct(station_name, line, ada) %>% filter(ada == "TRUE") %>% count()` are ADA compliant.

## Problem 3
#### clean the data in pols-month.csv.

```{r message = FALSE, echo = FALSE}
month_df = tibble(month = 1:12, month_name = month.name)
month_df
```

```{r message = FALSE, echo = FALSE}
#separate month to year, month, date
pols_df = read_csv(
  "./data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "date")) %>%
  mutate(month = as.integer(month)) 
```

```{r}
#replace month number with month name (need to delet month col by yourself.)
#create a president variable taking values gop and dem
#remove prez_dem and prez_gop; and remove the day variable.
pols_month_df = left_join(pols_df, month_df, by = "month") %>% select(-month) %>% 
  pivot_longer(
  cols = starts_with("prez_"),
    names_to = "president",
    names_prefix = "prez_gop_dem",
    values_to = "value") %>% 
  select(-date) %>% 
  relocate(year, month_name, president, value, everything()) 
```




